Self-interacting dark matter model without dark energy in cosmology

### 2024-11-29 article/article.tex
move the cmb power sepctrum methods

   If $n=0$, the global density evolution can be transformed as
   \begin{equation}
      \rho(a)=C_1+a^{-3}\frac{e^{-2\kappa C_1(t-t_0)}}{C_2-\kappa Z(t)},
   \end{equation}
   where $Z(t)=\int_{t_0}^ta^{-3}e^{-2\kappa C_1(t'-t_0)}\mathrm{d}t'$(\cite{refId0}).
   Using the usual unit-less $\Omega$-like parameterization, we obtain that
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{1,0}+\frac{\Omega_{1,0}\Omega_{2,0}a^{-3}e^{-2\kappa C_1(t-t_0)}}
      {\Omega_{1,0}+\kappa C_1\Omega_{2,0}Z(t)},
   \end{equation}
   where $\Omega_{1,0}=8\pi GC_1/3H_0^2$ and $\Omega_{2,0}=8\pi G/3H_0^2C_2$,
   which related to $\Omega_{\Lambda}$ and $\Omega_m$ in the $\Lambda$CDM model,
   respectively.
   From Eq.({\ref{eq:2}}), we can also give the mass of the DM particle
   related to the range of $\kappa C_1$(in the unit of Gyr${}^{-1}$)
   \begin{equation}
      M_x=\frac{3.325\times10^{-12}}{\kappa C_1}
      \frac{\langle\sigma u\rangle}{10^{-23}}h^2(1-\Omega_{2,0})\,\text{GeV},
   \end{equation}
   where $h\equiv H_0/[100\text{km/s/Mpc}]$.

@@

\section{Observational data}

   To constrain the relevant IDM models (\cite{refId0}), we use the newly revised
   observational $H(z)$ data (OHD)(\cite{Zhang_2014};\cite{PhysRevD.71.123001};
   \cite{Daniel.Stern_2010};\cite{M.Moresco_2012};\cite{Moresco_2016};
   \cite{10.1093/mnras/stx301};\cite{10.1093/mnrasl/slv037};\cite{Borghi_2022};
   \cite{Jiao_2023}),the Pantheon+ set of 1701 SNe Ia (), the CMB data from
   Planck 2018 and the BAO data from SDSS and DESI 2024.

@@

\subsection{The observational SNe Ia data}

   SNe Ia have long been used as "standard candles" to give a direct
   measurement of their luminosity distance, and provides strong constraints
   on cosmological parameters. We use the latest Pantheon+ data set of 1701 
   SNe Ia samples, which covers the redshift range [0, 2.26].
   \\We could use the simple likelihood function as
   \begin{equation}
      \chi_{S}^2=\sum_{i=1}^{1701}\frac{(\mu_{\text{th}}-\mu_{\text{data}})^2}{\sigma_i^2}.
   \end{equation}

\subsection{The baryon acoustic oscillation data}

>>>>

   If $n=0$, the global density evolution can be transformed as
   \begin{equation}
      \rho(a)=C_1+a^{-3}\frac{e^{-2\kappa C_1(t-t_0)}}{C_2-\kappa Z(t)},
   \end{equation}
   where $Z(t)=\int_{t_0}^ta^{-3}e^{-2\kappa C_1(t'-t_0)}\mathrm{d}t'$(\cite{refId0}).
   Using the usual unit-less $\Omega$-like parameterization, we obtain that
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{1,0}+\frac{\Omega_{1,0}\Omega_{2,0}a^{-3}e^{-2\kappa C_1(t-t_0)}}
      {\Omega_{1,0}+\kappa C_1\Omega_{2,0}Z(t)},
   \end{equation}
   where $\Omega_{1,0}=8\pi GC_1/3H_0^2$ and $\Omega_{2,0}=8\pi G/3H_0^2C_2$,
   which related to $\Omega_{\Lambda}$ and $\Omega_m$ in the $\Lambda$CDM model,
   respectively.
   From Eq.({\ref{eq:2}}), we can also give the mass of the DM particle
   related to the range of $\kappa C_1$(in the unit of Gyr${}^{-1}$)
   \begin{equation}
      M_x=\frac{3.325\times10^{-12}}{\kappa C_1}
      \frac{\langle\sigma u\rangle}{10^{-23}}h^2(1-\Omega_{2,0})\,\text{GeV},\label{eq:9}
   \end{equation}
   where $h\equiv H_0/[100\text{km/s/Mpc}]$.

@@

\section{Dataset}

   To constrain the relevant IDM models (\cite{refId0}), we use the newly revised
   observational $H(z)$ data (OHD)(\cite{Zhang_2014};\cite{PhysRevD.71.123001};
   \cite{Daniel.Stern_2010};\cite{M.Moresco_2012};\cite{Moresco_2016};
   \cite{10.1093/mnras/stx301};\cite{10.1093/mnrasl/slv037};\cite{Borghi_2022};
   \cite{Jiao_2023}),the Pantheon+ set of 1701 SNe Ia (\cite{Scolnic_2022}), the CMB data from
   Planck 2018 and the BAO data from SDSS and DESI 2024.

@@

\subsection{Type Ia supernovae}

   SNe Ia have long been used as "standard candles" to give a direct
   measurement of their luminosity distance, and provides strong constraints
   on cosmological parameters. We use the latest Pantheon+ data set of 1701 
   SNe Ia samples(\cite{Scolnic_2022}), which covers the redshift range [0, 2.26].

   We use the fiducial SN Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances(\cite{Riess_2022}), which gives the $\mu_\text{data}$ and constrains $H_0$
   in advance. To eliminate the influence of $M_b$, we give the likelihood function
   as \begin{equation}
      \chi_{\text{S}}^2=A-\frac{B^2}{C}+\ln\left(\frac{C}{2\pi}\right),
   \end{equation}
   where $A=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})^2/\sigma_i^2,
   B=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})/\sigma_i^2,C=\sum_i^{1701}1/\sigma_i^2$,
   the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.
   \end{equation}
   However, the Eq.(\ref{eq:9}) just need the $H_0$ to caculate the $M_x$
   and we would still use the simple likelihood function as
   \begin{equation}
      \chi_{\text{SNe}}^2=A=\sum_{i=1}^{1701}\frac{(\mu_{\text{th}}-\mu_{\text{data}})^2}{\sigma_i^2}.
   \end{equation}
   Apparently these two functions give the same constraints in $\Omega_{2,0}$
   and $\kappa C_1$.

\subsection{Baryon acoustic oscillation}

   The Baryon acoustic oscillation method (BAO) provides a key cosmological probe
   sensitive to the cosmic expansion history with well-controlled systematics.
   We use two BAO data sets from the SDSS(\cite{PhysRevD.103.083533}) and DESI 2024(\cite{desicollaboration2024desi2024vicosmological}),
   which are given at Table \ref{tab:2} and Table \ref{tab:3}, respectively.

   \begin{table}[htbp]
      \caption{The BAO-only dataset from SDSS}
      \centering
      \begin{tabular}{lllll}
         \hline\hline
            $z_{\text{eff}}$ & $D_{\text{M}}/r_{\text{d}}$ & $D_{\text{H}}/r_{\text{d}}$ & $D_{\text{V}}/r_{\text{d}}$ \\
            \hline
            0.15 & & & 4.47$\pm$0.17 \\
            0.38 & 10.23$\pm$0.17 & 25$\pm$0.76 & \\
            0.51 & 13.36$\pm$0.21 & 22.33$\pm$0.58 & \\
            0.7 & 17.86$\pm$0.33 & 19.33$\pm$0.53 & \\
            0.85 & & & $18.33_{-0.62}^{+0.57}$ \\
            1.48 & 30.69$\pm$0.8 & 13.26$\pm$0.55 & \\
            2.33 & 37.6$\pm$1.9 & 8.93$\pm$0.28 & \\
            2.33 & 37.3$\pm$1.7 & 9.08$\pm$0.34 & \\
            \hline
      \end{tabular}
      \label{tab:2}
   \end{table}

   \begin{table}[htbp]
      \caption{The BAO dataset from DESI 2024}
      \centering
      \begin{tabular}{lllll}
         \hline\hline
            $z_{\text{eff}}$ & $D_{\text{M}}/r_{\text{d}}$ & $D_{\text{H}}/r_{\text{d}}$ & $D_{\text{V}}/r_{\text{d}}$ \\
            \hline
            0.295 & & & 7.93$\pm$0.15 \\
            0.51 & 13.62$\pm$0.25 & 20.98$\pm$0.61 & \\
            0.706 & 16.85$\pm$0.32 & 20.08$\pm$0.6 & \\
            0.93 & 21.71$\pm$0.28 & 17.88$\pm$0.35 & \\
            1.317 & 27.79$\pm$0.69 & 13.82$\pm$0.42 & \\
            1.491 & & & 26.07$\pm$0.67 \\
            2.33 & 39.71$\pm$0.94 & 8.52$\pm$0.17 & \\
            \hline
      \end{tabular}
      \label{tab:3}
   \end{table}

\subsection{Cosmic Microwave Background}

   We adopt the $\Omega_bh^2$ from BBN,
   give the constraint to $r_d$
